---
title: "Manuscript Figures"
author: "Javiera Rudolph"
date: "March 25, 2019"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggtern)
```


```{r Functions}
# FUNCTIONS ---------------------------------------------------------------


# ** get_sites_data() to organize site level data and get ready for plot

    # This function takes the simulated metacommunity to get the species richness for each site
    # It then binds that information to the Variation Partinioning output from the HMSC::variPart()
    # This function also takes the output of HMSC::variPart() and organizes each fraction:
    #   into the env, spa and codist components
    # The output of the get_sites_data() function is a dataframe that is ready to plot

get_sites_data <- function(folderpath, scenario){
  richness <- readRDS(paste0(folderpath, scenario, "-metacomSim.RDS")) %>% 
    set_names(imap(., ~ paste0("iter", .y))) %>% 
    map(., rowSums) %>%
    bind_rows() %>% 
    rownames_to_column(var = "sites") %>% 
    gather(., key = "iteration", value = "richness", -sites) %>% 
    mutate(identifier = paste0("site", sites, "_", iteration)) %>% 
    dplyr::select(., -c(sites, iteration))
  

  vp <- readRDS(paste0(folderpath, scenario, "-vpsites.RDS"))
  
  overlap1 <- map(vp, "overlap1")
  overlap2 <- map(vp, "overlap2")
  overlap3 <- map(vp, "overlap3")
  
  vpALL <- vector("list", length = 5)
  for(i in 1:5){
    workingVP1 <- overlap1[[i]]
    workingVP2 <- overlap2[[i]]
    workingVP3 <- overlap3[[i]]
    
    c <- rowSums(workingVP1[,,1])/15
    b <- rowSums(workingVP1[,,2])/15
    a <- rowSums(workingVP1[,,3])/15
    
    e <- rowSums(workingVP2[,,1])/15
    f <- rowSums(workingVP2[,,2])/15
    d <- rowSums(workingVP2[,,3])/15
    
    g <- rowSums(workingVP3)/15
    
    env <- a + f + 1/2 * d + 1/2 * g
    env <- ifelse(env < 0, 0, env)
    spa <- b + e + 1/2 * d + 1/2 * g
    spa <- ifelse(spa < 0, 0, spa)
    random <- c
    codist <- ifelse(random < 0, 0, random)
    r2 <- env + spa + codist
    iteration <- factor(paste0("iter", i), levels = paste0("iter", 1:5))
    
    cleanData <- cbind.data.frame(env, spa, codist, r2, iteration)
    cleanData$site <- paste0(row.names(cleanData))
    
    vpALL[[i]] <- cleanData
  }
  
  vpALL %>% 
    bind_rows() %>% 
    mutate(identifier = paste(site, iteration, sep = "_"),
           scenario = scenario) %>% 
    left_join(., richness)
}


# ** get_species_data() this function organizes species level data and gets it ready to plot.
  # It takes the RDS file that is the output of the HMSC::variPart() function and reorganizes it into the 
  # env, codist and spa components.
  # It also calculates the prevalence information for each species. 

get_species_data <- function(folderpath, scenario){
  
  prevalence <- readRDS(paste0(folderpath, scenario, "-metacomSim.RDS")) %>% 
    set_names(imap(., ~ paste0("iter_", .y))) %>% 
    map(., colSums) %>%
    bind_cols() %>% 
    rownames_to_column(var = "species") %>% 
    gather(., key = "iteration", value = "prevalence", -species) %>% 
    mutate(identifier = paste0("spp", species, "_", iteration),
           prevalence = prevalence/1000) %>% 
    dplyr::select(., -c(species, iteration))
  
  
  
  
  readRDS(paste0(folderpath, scenario, "-vpspp.RDS")) %>% 
    set_names(imap(., ~ paste0("iter_", .y))) -> VPdata
  
  fullData <- list()
  for(i in 1:length(VPdata)){
    fullData[[i]] <- VPdata[[i]] %>% 
      map(as_tibble) %>%
      bind_cols() %>% 
      rownames_to_column() %>% 
      set_names(c("species", "c", "b", "a", "e", "f", "d", "g")) %>% 
      transmute(species = species,
                env = a + f + 0.5 * d + 0.5 * g,
                env = ifelse(env < 0, 0, env),
                spa = b + e + 0.5 * d + 0.5 * g,
                spa = ifelse(spa < 0, 0, spa),
                codist = c,
                codist = ifelse(codist < 0, 0, codist),
                r2 = env + spa + codist,
                iteration = names(VPdata[i]))
    
  }
  
  fullData %>% 
    bind_rows() %>% 
    mutate(identifier = paste0("spp", species, "_", iteration), 
           scenario = scenario) %>% 
    left_join(., prevalence)
  
}


# ** get_fig2_params() This function is for Figure 2 only. 
# It takes the RDS params file and organizes it from a list to a dataframe, so that it can be joined to the
# Variation Partition results at the species level. That way we can know what niche values each species has.
# The output of this function can be joined to the VP results at the species level and then plot

get_fig2_params <- function(folderpath, scenario){
  pars <- readRDS(paste0(folderpath, scenario, "-params.RDS"))
  
  with(pars, {enframe(u_c[1,], name = "species", value = "nicheOptima") %>% 
      left_join(., enframe(s_c[1,], name = "species", value = "nicheBreadth")) %>% 
      left_join(., enframe(c_0, name = "species", value = "colonizationProb")) %>% 
      mutate(., dispersal = alpha,
             species = as.character(species),
             speciesChr = as.character(paste0("spp_", species)),
             interCol = d_c ,
             interExt = d_e)})
}

# ** get_fig3_params() this is for Figure 3 only. 
# Figure 3 has three sets of parameters because we divided simulations into groups.
# Half of the species in each simulation has interactions, and the other half doesn't.
# Or, a third of the species is assigned a different dispersal level, etc...

get_fig3_params <- function(folderpath, scenario){
  parsList <- readRDS(paste0(folderpath, scenario, "-params.RDS"))
  fullPars <-  data.frame()
  for(i in 1:length(parsList)){
    
    nspp <- nrow(fullPars)
    
    i_pars <- with(parsList[[i]], {
      enframe(u_c[1,], name = "species",value =  "nicheOpt") %>% 
        left_join(., enframe(s_c[1,], name = "species", value =  "nicheBreadth")) %>% 
        left_join(., enframe(c_0, name = "species", value =  "colProb")) %>% 
        mutate(dispersal = as.factor(alpha), 
               species = as.character(species + nspp), 
               intercol = as.factor(d_c), 
               interext = d_e)})
    
    
    fullPars <- rbind(fullPars, i_pars)
    
  }
  
  return(fullPars)
}
```

```{r PlotFunctions}

# ** species_plot() it generates the species level plot, coloring by iteration
species_plot <- function(data, scenario, colorVar = NULL, colorLegend = "none"){
  data %>% 
    ggtern(aes(x = env, z = spa, y = codist, size = r2)) +
  geom_point(aes_string(color = colorVar), alpha = 0.8) +
  scale_T_continuous(limits=c(0,1.0),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  scale_L_continuous(limits=c(0.0,1),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  scale_R_continuous(limits=c(0.0,1.0),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  labs(title = paste(scenario, "- Species"),
       x = "E",
       xarrow = "Environment",
       y = "C",
       yarrow = "Co-Distribution",
       z = "S", 
       zarrow = "Spatial Autocorrelation") +
  theme_light() +
  theme_showarrows() +
  #scale_colour_brewer(palette = "Set1") +
  #scale_colour_brewer(palette = "Spectral") +
  #scale_color_viridis_d() +
  scale_size_area(limits = c(0,1), breaks = seq(0,1,0.2)) +
  guides(color = guide_legend(colorLegend, order = 2), 
         size = guide_legend(title = expression(R^2), order = 1)) +
  theme(panel.grid = element_line(color = "darkgrey"),
        axis.title = element_text(size = 8))
}

# Plot sites by dots
sites_plot <- function(data, scenario, colorVar = NULL, colorLegend = "none"){
  data %>% 
    ggtern(aes(x = env, z = spa, y = codist, size = r2)) +
  geom_point(aes_string(color = colorVar), alpha = 0.6) +
  scale_T_continuous(limits=c(0,1.0),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  scale_L_continuous(limits=c(0.0,1),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  scale_R_continuous(limits=c(0.0,1.0),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  labs(title = paste(scenario, "- Sites"),
       x = "E",
       xarrow = "Environment",
       y = "C",
       yarrow = "Co-Distribution",
       z = "S", 
       zarrow = "Spatial Autocorrelation") +
  theme_light() +
  theme_showarrows() +
  #scale_colour_brewer(palette = "Set1") +
  #scale_colour_brewer(palette = "Spectral") +
  #scale_color_viridis_d() +
  scale_size_area(limits = c(0, 0.003), breaks = seq(0, 0.003, 0.0005)) +
  guides(color = guide_legend(colorLegend), 
         size = guide_legend(title = expression(R^2))) +
  theme(panel.grid = element_line(color = "darkgrey"),
        axis.title = element_text(size = 8))
}

sites_dens_plot <- function(data, scenario){
  data %>% 
    ggtern(aes(x = env, z = spa, y = codist)) +
  stat_density_tern(geom = "polygon", aes(alpha = ..level.., fill = ..level..))+
  theme_minimal() +
  scale_T_continuous(limits=c(0,1.0),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  scale_L_continuous(limits=c(0.0,1),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  scale_R_continuous(limits=c(0.0,1.0),
                     breaks=seq(0,1,by=0.1),
                     labels=seq(0,1,by=0.1)) +
  labs(title = paste(scenario, "- Sites"),
       x = "E",
       xarrow = "Environment",
       y = "C",
       yarrow = "Co-Distribution",
       z = "S",
       zarrow = "Spatial Autocorrelation") +
  theme_light() +
  theme_showarrows() +
  #scale_fill_gradient(low = "yellow", high = "red") +
  #scale_fill_gradient(los = "blue", high = "orange") +
  guides(fill = guide_legend(title = "Density"),
         alpha = guide_legend("none")) +
  theme(panel.grid = element_line(color = "darkgrey"),
        axis.title = element_text(size = 8))
}
```


```{r}
folderpath <- "outputs/MANUSCRIPT/"
fig2Scenarios <- c("FIG2A", "FIG2B", "FIG2C", "FIG2D")

```

All of these results come from using a quadratic response to the environment

# Figure 2

## Option1
All figures are color coded by iteration. The $R^2$ scales are shared by figures at the species level and a separate scale is shared at the sites level. 

```{r fig.width=14, fig.height=6}

for(i in 1:4){
  scenario <- fig2Scenarios[i]
  
  sppData <- get_species_data(folderpath, scenario)
  sppPlot <- species_plot(sppData, scenario, colorVar = "iteration") + 
    scale_colour_brewer(palette = "Spectral")
  
  sppPlot
  sitesData <- get_sites_data(folderpath, scenario)
  sitesPlot <- sites_plot(sitesData, scenario, colorVar = "iteration") + 
    scale_colour_brewer(palette = "Spectral")
  sitesPlot
  
  grid.arrange(arrangeGrob(sppPlot),
               arrangeGrob(sitesPlot), ncol = 2)
}

```

## Option 2  

Same results, but using a density plot for the sites level data. A higher "density" of sites is represented by the lighter color on the scale. 
```{r fig.width=14, fig.height=6}
for(i in 1:4){
  scenario <- fig2Scenarios[i]
  sppData <- get_species_data(folderpath, scenario)
  sppPlot <- species_plot(sppData, scenario, colorVar = "iteration") +
    scale_colour_viridis_d()
  
  sitesData <- get_sites_data(folderpath, scenario)
  sitesPlot <- sites_dens_plot(sitesData, scenario) +
    scale_fill_viridis_c()
  
  grid.arrange(arrangeGrob(sppPlot),
               arrangeGrob(sitesPlot), ncol = 2)
}

```

#Figure 3  


## Option1

```{r fig.height=6, fig.width=14}
fig3Scenarios <- c("FIG3A", "FIG3B", "FIG3C")
colorVar <- c("intercol", "dispersal", "dispersal")
colLegend <- c("Interaction \n level", "Dispersal \n level", "Dispersal \n level")

for(i in 1:3){
  scenario <- fig3Scenarios[i]
  
  sppData <- get_species_data(folderpath, scenario) %>% 
    left_join(., get_fig3_params(folderpath, scenario))
  sppPlot <- species_plot(sppData, scenario, colorVar = colorVar[i], colorLegend = colLegend[i]) + 
    scale_colour_brewer(palette = "Set1")
  
  sitesData <- get_sites_data(folderpath, scenario)
  sitesPlot <- sites_plot(sitesData, scenario, colorVar = "iteration") +
    scale_colour_brewer(palette = "Spectral")
  
  grid.arrange(arrangeGrob(sppPlot),
               arrangeGrob(sitesPlot), ncol = 2)
}
```


## Option 2


```{r fig.width=14, fig.height=6}
for(i in 1:3){
  scenario <- fig3Scenarios[i]
  
  sppData <- get_species_data(folderpath, scenario) %>% 
    left_join(., get_fig3_params(folderpath, scenario))
  sppPlot <- species_plot(sppData, scenario, colorVar = colorVar[i], colorLegend = colLegend[i]) +
    scale_colour_viridis_d(option = "magma")
  
  sitesData <- get_sites_data(folderpath, scenario)
  sitesPlot <- sites_dens_plot(sitesData, scenario) +
    scale_fill_viridis_c(option = "magma")
  
  grid.arrange(arrangeGrob(sppPlot),
               arrangeGrob(sitesPlot), ncol = 2)
}

```