---
title: "Figure 3 - Comparisons between gaussian and quadratic response to the environment"
author: "Javiera Rudolph"
date: "March 13, 2019"
output: github_document
---


The results in this document correspond to the comparisons between gaussian and quadratic responses to the environment. We realized that using the same niche breadth in the gaussian and quadratic responses was dramatically different. The area under the curve of the parabola is significantly smaller, making the response to the environment extremely restricted. Therefore, we modified the parameters so that we could mantain a similar response when using a quadratic and gaussian response. Specifically, the quadratic function uses a larger niche breadth so keep a similar area under the curve as the gaussian response. When editing this function (S_f: response to environment), using a parabola would give us negative values for areas outside the range of occurence. This might influence interactions, so we decided to rerun the scripts with all negative values in the response to environment set equal to zero. This is shown in the third figure of each scenario. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(ggtern)
```

```{r functions}

get_params_table <- function(folderpath, scenario){
  parsList <- readRDS(paste0(folderpath, scenario, "-params.RDS"))
  fullPars <-  data.frame()
  for(i in 1:length(parsList)){
    
    nspp <- nrow(fullPars)
    
    i_pars <- with(parsList[[i]], {
      enframe(u_c[1,], name = "species",value =  "nicheOpt") %>% 
        left_join(., enframe(s_c[1,], name = "species", value =  "nicheBreadth")) %>% 
        left_join(., enframe(c_0, name = "species", value =  "colProb")) %>% 
        mutate(dispersal = alpha, 
               species = as.character(species + nspp), 
               intercol = d_c, 
               interext = d_e)})
    
    
    fullPars <- rbind(fullPars, i_pars)
    
  }
  fullPars
}


organize_vp_overlaps <- function(folderpath, scenario, indSites = FALSE){
  
  
  if(indSites == TRUE){
    richness <- readRDS(paste0(folderpath, scenario, "-metacomSim.RDS")) %>% 
      set_names(imap(., ~ paste0("iter", .y))) %>% 
      map(., rowSums) %>%
      bind_rows() %>% 
      rownames_to_column(var = "sites") %>% 
      gather(., key = "iteration", value = "richness", -sites) %>% 
      mutate(identifier = paste0("site", sites, "_", iteration)) %>% 
      select(., -c(sites, iteration))
    
    
    
    vp <- readRDS(paste0(folderpath, scenario, "-vpsites.RDS"))
    
    overlap1 <- map(vp, "overlap1")
    overlap2 <- map(vp, "overlap2")
    overlap3 <- map(vp, "overlap3")
    
    vpALL <- vector("list", length = 5)
    for(i in 1:5){
      workingVP1 <- overlap1[[i]]
      workingVP2 <- overlap2[[i]]
      workingVP3 <- overlap3[[i]]
      
      c <- rowSums(workingVP1[,,1])/15
      b <- rowSums(workingVP1[,,2])/15
      a <- rowSums(workingVP1[,,3])/15
      
      e <- rowSums(workingVP2[,,1])/15
      f <- rowSums(workingVP2[,,2])/15
      d <- rowSums(workingVP2[,,3])/15
      
      g <- rowSums(workingVP3)/15
      
      env <- a + f + 1/2 * d + 1/2 * g
      env <- ifelse(env < 0, 0, env)
      spa <- b + e + 1/2 * d + 1/2 * g
      spa <- ifelse(spa < 0, 0, spa)
      random <- c
      codist <- ifelse(random < 0, 0, random)
      r2 <- env + spa + codist
      iteration <- factor(paste0("iter", i), levels = paste0("iter", 1:5))
      
      cleanData <- cbind.data.frame(env, spa, codist, r2, iteration)
      cleanData$site <- paste0(row.names(cleanData))
      
      vpALL[[i]] <- cleanData
    }
    
    vpALL %>% 
      bind_rows() %>% 
      mutate(identifier = paste(site, iteration, sep = "_"),
             scenario = scenario) %>% 
      left_join(., richness) -> vpALL
    
    return(vpALL)
  }
  
  parsList <- readRDS(paste0(folderpath, scenario, "-params.RDS"))
  fullPars <-  data.frame()
  for(i in 1:length(parsList)){
    
    nspp <- nrow(fullPars)
    
    i_pars <- with(parsList[[i]], {
      enframe(u_c[1,], name = "species",value =  "nicheOpt") %>% 
        left_join(., enframe(s_c[1,], name = "species", value =  "nicheBreadth")) %>% 
        left_join(., enframe(c_0, name = "species", value =  "colProb")) %>% 
        mutate(dispersal = alpha, 
               species = as.character(species + nspp), 
               intercol = d_c, 
               interext = d_e)})
    
    
    fullPars <- rbind(fullPars, i_pars)
    
  }
  
  params <- fullPars
  
  prevalence <- readRDS(paste0(folderpath, scenario, "-metacomSim.RDS")) %>% 
    set_names(imap(., ~ paste0("iter_", .y))) %>% 
    map(., colSums) %>%
    bind_cols() %>% 
    rownames_to_column(var = "species") %>% 
    gather(., key = "iteration", value = "prevalence", -species) %>% 
    mutate(identifier = paste0("spp", species, "_", iteration)) %>% 
    select(., -c(species, iteration))
  
  
  
  
  readRDS(paste0(folderpath, scenario, "-vpspp.RDS")) %>% 
    set_names(imap(., ~ paste0("iter_", .y))) -> VPdata
  
  fullData <- list()
  for(i in 1:length(VPdata)){
    fullData[[i]] <- VPdata[[i]] %>% 
      map(as_tibble) %>%
      bind_cols() %>% 
      rownames_to_column() %>% 
      set_names(c("species", "c", "b", "a", "e", "f", "d", "g")) %>% 
      transmute(species = species,
                env = a + f + 0.5 * d + 0.5 * g,
                env = ifelse(env < 0, 0, env),
                spa = b + e + 0.5 * d + 0.5 * g,
                spa = ifelse(spa < 0, 0, spa),
                codist = c,
                codist = ifelse(codist < 0, 0, codist),
                r2 = env + spa + codist,
                iteration = names(VPdata[i])) %>% 
      left_join(., params)
    
  }
  
  fullData %>% 
    bind_rows() %>% 
    mutate(identifier = paste0("spp", species, "_", iteration), 
           scenario = scenario) %>% 
    left_join(., prevalence) -> fullData
  return(fullData)
  
}


make_tern_plot <- function(figData, varShape = NULL, varColor = NULL){
  figData %>% 
    ggtern(aes(x = env, z = spa, y = codist, size = r2)) +
    scale_T_continuous(limits=c(0,1.0),
                       breaks=seq(0,1,by=0.1),
                       labels=seq(0,1,by=0.1)) +
    scale_L_continuous(limits=c(0.0,1),
                       breaks=seq(0,1,by=0.1),
                       labels=seq(0,1,by=0.1)) +
    scale_R_continuous(limits=c(0.0,1.0),
                       breaks=seq(0,1,by=0.1),
                       labels=seq(0,1,by=0.1)) +
    labs(x = "Environment",
         y = "Codistribution",
         z = "Spatial") +
    geom_point(aes_string(shape = varShape, color = varColor), alpha = 0.7) +
    theme_minimal() +
    guides(size = guide_legend(order = 1,
                               title = expression(paste(R^{2}))),
           shape = guide_legend("none"),
           # shape = guide_legend(order = 2,
           #                      title = varShape,
           #                      override.aes = list(size = 3)),
           col = guide_colourbar(title = varColor,
                                 order = 3)) +
    theme(#legend.position = "bottom", legend.box = "vertical",
      panel.grid = element_line(color = "darkgrey"),
      axis.title = element_text(size = 7))
}
  

# Arrange plots -----------------------------------------------------------
arrange_plots <- function(folderpath, scenario, sppColor){
  organize_vp_overlaps(folderpath, scenario, indSites = FALSE) %>% 
    make_tern_plot(., varShape = "iteration", varColor = sppColor) -> sppPlot
  
  organize_vp_overlaps(folderpath, scenario, indSites = TRUE) %>% 
  make_tern_plot(., varShape = "iteration", varColor = "richness") -> sitesPlot
  
  grid.arrange(arrangeGrob(sppPlot),
               arrangeGrob(sitesPlot),
               ncol = 2)
}


```

```{r folderPaths}
gaussianPath <- "outputs/20190312-fifteen_species_fig3_gaussian/"
quadraticPath <- "outputs/20190313-fifteen_species_fig3_quadratic/"
quadraticPathNoNegs <- "outputs/20190314-fifteen_species_fig3_quadratic_nonegs/"
```


# Scenario 1
# Scenario 1 - Gaussian response

```{r}
scenario <- "scenario1"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)[c(1,8),3:7]
knitr::kable(parstable)
```


```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario, sppColor = "intercol")
```


# Scenario 1 - Quadratic response
```{r}
folderpath <- quadraticPath
parstable <- get_params_table(folderpath, scenario)[c(1,8),3:7]
knitr::kable(parstable)
```

```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario, sppColor = "intercol")
```

### No negative values in the response to environment

```{r, fig.width=10, fig.height=6}

arrange_plots(quadraticPathNoNegs, scenario, sppColor = "intercol")
```



# Scenario 2
# Scenario 2 - Gaussian response

```{r}
scenario <- "scenario2"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)[c(1,6,11),3:7]
knitr::kable(parstable)
```


```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario, sppColor = "dispersal")
```


# Scenario 2 - Quadratic response
```{r}
folderpath <- quadraticPath
parstable <- get_params_table(folderpath, scenario)[c(1,6,11),3:7]
knitr::kable(parstable)
```

```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario, sppColor = "dispersal")
```

### No negative values in the response to environment

```{r, fig.width=10, fig.height=6}

arrange_plots(quadraticPathNoNegs, scenario, sppColor = "dispersal")
```



# Scenario 3
# Scenario 3 - Gaussian response

```{r}
scenario <- "scenario3"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)[c(1,6,11),3:7]
knitr::kable(parstable)
```


```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario, sppColor = "dispersal")
```


# Scenario 3 - Quadratic response
```{r}
folderpath <- quadraticPath
parstable <- get_params_table(folderpath, scenario)[c(1,6,11),3:7]
knitr::kable(parstable)
```

```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario, sppColor = "dispersal")
```

### No negative values in the response to environment

```{r, fig.width=10, fig.height=6}

arrange_plots(quadraticPathNoNegs, scenario, sppColor = "dispersal")
```


```{r save csvFilesMathew, eval=FALSE}

scen3spp <- organize_vp_overlaps(folderpath, scenario, indSites = FALSE)
write.csv(scen3spp, file = paste0(folderpath, "csvFiles/scen3spp.csv"))

scen3sites <- organize_vp_overlaps(folderpath, scenario, indSites = TRUE)
write.csv(scen3spp, file = paste0(folderpath, "csvFiles/scen3sites.csv"))

```








