---
title: "Species by Species covariance (or correlation) matrix"
author: "Javiera Rudolph"
date: "March 21, 2019"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

library(tidyverse)
library(ggtern)
library(HMSC)
library(corrplot)
library(circlize)
```

```{r functions}
source("functions/rmd_figure_prep_fx.R")

```

We would like to get the species to species interaction matrices for each of the scenarios being tested. When setting up the metacommunity simulations, an interaction matrix, **A**, was used. We would like to compare the interaction matrix resulting from the model to the one used in the simulation. From the original code, this matrix **A** is defined as:

```{r matrixA}
# Interaction matrix
R <- 15
A = matrix(0,nr=R,nc=R)
d = as.matrix(dist(c(1:R),upper=TRUE,diag=T))
A[d<=1] = -1
diag(A) = 0

A

```

Using the example from scenario in Figure 2c, fow which we used the following simulation parameters:

```{r}
folderpath <- "outputs/20190318-fig2c_quad/"
params <- get_params_table(folderpath, "scenario4")
params %>% 
  knitr::kable()
```

And we obtained the following results:

```{r fig.width=14, fig.height=4}
arrange_plots(folderpath, "scenario4", colorSpp = "nicheOpt", colorSites = "richness", leftlabel = "Fig2c")
```

When we look at the structure of the HMSC object created for this scenario, it has 5 iterations. In the resulting figures, each iteration has a different shape. Does this mean that we should get a species interaction matrix for each one of these iterations? 
```{r}
modelfile <- readRDS(paste0(folderpath, "scenario4-model.RDS"))
```

```{r, fig.height=14, fig.width=14}
par(mfrow=c(3,2))

for(i in 1:5){
assoMat <- corRandomEff(modelfile[[i]])
chordDiagram(assoMat[, , 1, 1], symmetric = TRUE)
}

```

When we want to look at the matrix, I am confused about the siteMean vs plotMean. Keep in mind that this site mean is using only one of the iterations. 

```{r}
for(i in 1:5){
  assoMat <- corRandomEff(modelfile[[i]])
siteMean <- apply(assoMat[ , , , 1], 1:2, mean)
#plotMean <- apply(assoMat[ , , , 2], 1:2, mean)

siteDrawCol <- matrix(NA, nrow = nrow(siteMean), ncol = ncol(siteMean))
siteDrawCol[which(siteMean > 0.4, arr.ind=TRUE)]<-"red"
siteDrawCol[which(siteMean < -0.4, arr.ind=TRUE)]<-"blue"
# Build matrix of "significance" for corrplot
siteDraw <- siteDrawCol
siteDraw[which(!is.na(siteDraw), arr.ind = TRUE)] <- 0
siteDraw[which(is.na(siteDraw), arr.ind = TRUE)] <- 1
siteDraw <- matrix(as.numeric(siteDraw), nrow = nrow(siteMean), ncol = ncol(siteMean))

par(mfrow=c(1,2))
Colour <- colorRampPalette(c("blue", "white", "red"))(200)
corrplot(siteMean, method = "color", col = Colour, type = "lower",
diag = FALSE, p.mat = siteDraw, tl.srt = 45)

corrplot(A, method = "color", col = Colour, type = "lower",
diag = FALSE, p.mat = siteDraw, tl.srt = 45)
}
```

