---
title: "Two species - when does niche breadth allow interaction?"
author: "Javiera Rudolph"
date: "March 18, 2019"
output: github_document
---

We tried changing the niche breadth (increase) to allow for interactions to play a role in the Figure2c scenario. It didn't work so now we are trying to test a range of niche breadths to see at which point we are able to see interactions. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(ggtern)
```

```{r functions, include=FALSE}

source("functions/data_wrangling_fx.R")
source("functions/make_tern_plot_fx.R")

arrange_plots <- function(folderpath, scenario){
  doItAll_dataWrangling(outPath = folderpath, scenarioNum = scenario, indSites = FALSE) %>% 
    make_tern_plot(., varShape = "iteration", varColor = "nicheOpt") -> sppPlot
  
  sitescsv <- doItAll_dataWrangling(outPath = folderpath, scenarioNum = scenario, indSites = TRUE)
  make_tern_plot(sitescsv, varShape = "iteration", varColor = "richness") -> sitesPlot
  
  grid.arrange(arrangeGrob(sppPlot),
               arrangeGrob(sitesPlot),
               ncol = 2)
}

get_params_table <- function(folderpath, scenario){
  params <- with(readRDS(paste0(folderpath, scenario, "-params.RDS")), {
    enframe(u_c[1,], name = "species",value =  "nicheOpt") %>% 
      left_join(., enframe(s_c[1,], name = "species", value =  "nicheBreadth")) %>% 
      left_join(., enframe(c_0, name = "species", value =  "colProb")) %>% 
      mutate(dispersal = alpha, 
             species = as.character(species), 
             intercol = d_c, 
             interext = d_e)
  })
  
}

```

```{r setFolderPaths}

fig2c <- "outputs/20190317-fig2c_quad/"
fig2c_nonegs <- "outputs/20190318-fig2c_quad_nonegs/"

```




# Scenario1
```{r}
scenario <- "scenario1"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario2
```{r}
scenario <- "scenario2"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```


# Scenario3
```{r}
scenario <- "scenario3"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```


# Scenario4
```{r}
scenario <- "scenario4"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario5
```{r}
scenario <- "scenario5"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario6
```{r}
scenario <- "scenario6"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario7
```{r}
scenario <- "scenario7"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario8
```{r}
scenario <- "scenario8"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario9
```{r}
scenario <- "scenario9"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```

# Scenario10
```{r}
scenario <- "scenario10"
get_params_table(fig2c, scenario) %>% knitr::kable()
```


```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c, scenario)
```

No negatives 

```{r, fig.width=10, fig.height=6}
arrange_plots(fig2c_nonegs, scenario)
```
