---
title: "Two species niche comparisons between gaussian and quadratic response to environment"
author: "Javiera Rudolph"
date: "March 12, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(ggtern)
```


```{r setFolderPaths}


source("functions/rmd_figure_prep_fx.R")
gaussianPath <- "outputs/20190304-two_spp_niche_comparisons/"
quadraticPath <- "outputs/20190309-two_spp_niche_comparisons_quadratic/"


```

# Scenario 1 - Parameters
```{r}
scenario <- "scenario1"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 1 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 1 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```



# Scenario 2 - Parameters
```{r}
scenario <- "scenario2"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 2 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 2 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```



# Scenario 3 - Parameters
```{r}
scenario <- "scenario3"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 3 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 3 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```

# Scenario 4 - Parameters
```{r}
scenario <- "scenario4"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 4 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 4 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```


# Scenario 5 - Parameters
```{r}
scenario <- "scenario5"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 5 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 5 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```



# Scenario 6 - Parameters
```{r}
scenario <- "scenario6"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 6 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 6 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```





# Scenario 7 - Parameters
```{r}
scenario <- "scenario7"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 7 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 7 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```



# Scenario 8 - Parameters
```{r}
scenario <- "scenario8"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 8 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 8 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```



# Scenario 9 - Parameters
```{r}
scenario <- "scenario9"
folderpath <- gaussianPath
parstable <- get_params_table(folderpath, scenario)
knitr::kable(parstable)
```


# Scenario 9 - Gaussian response
```{r, fig.width=10, fig.height=6}

arrange_plots(folderpath, scenario)
```


# Scenario 9 - Quadratic response

```{r, fig.width=10, fig.height=6}
folderpath <- quadraticPath
arrange_plots(folderpath, scenario)
```